```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(mosaic)
library(car)
library(broom)
library(tidyverse)
library(lme4)
library(readxl)
library(writexl)
```

```{r}
inequality = read_excel("Data.xlsx")
inequality <- transform(inequality,GII = as.numeric(GII))
inequality <- transform(inequality,Gini = as.numeric(Gini))
inequality <- transform(inequality,HDI.group = as.factor(HDI.group))
drop_na(inequality)
```
#Describing variables
```{r}
str(inequality)
```
##GII:
```{r}
GII.summary = favstats(~ GII|HDI.group, data = inequality)
dotPlot(~ GII|HDI.group, data = inequality, main = "Dotplot of GII", xlab = "GII")
histogram(~ GII|HDI.group, data = inequality, main = "Histogram of GII", xlab = "GII", type = "count")
bwplot(~ GII|HDI.group, data = inequality, main = "Boxplot of GII", xlab = "GII", outer = TRUE)
```
##HDI:
```{r}
HDI.summary = favstats(~ HDI, data = inequality)
dotPlot(~ HDI, data = inequality, main = "Dotplot of HDI", xlab = "HDI")
histogram(~ HDI, data = inequality, main = "Histogram of HDI", xlab = "HDI", type = "count")
bwplot(~ HDI, data = inequality, main = "Boxplot of HDI", xlab = "HDI", outer = TRUE)
```
##Gini coefficient:
```{r}
Gini.summary = favstats(~ Gini|HDI.group, data = inequality)
dotPlot(~ Gini|HDI.group, data = inequality, main = "Dotplot of Gini", xlab = "Gini")
histogram(~ Gini|HDI.group, data = inequality, main = "Histogram of Gini", xlab = "Gini", type = "count")
bwplot(~ Gini|HDI.group, data = inequality, main = "Boxplot of Gini", xlab = "Gini", outer = TRUE)
```
##Relationship
```{r}
xyplot(GII ~ Gini, data = inequality, main = "Scatterplot between GII and Gini", xlab = "Gini", ylab = "GII", type = c("p","r"))

xyplot(GII ~ HDI, data = inequality, main = "Scatterplot between GII and HDI", xlab = "HDI", ylab = "GII", type = c("p","r"))

xyplot(Gini ~ HDI, data = inequality, main = "Scatterplot between Gini and HDI", xlab = "HDI", ylab = "Gini", type = c("p","r"))

scatterplotMatrix(~ GII + HDI + Gini|HDI.group, data = inequality)
```
###correlation
```{r}
cor.test(GII ~ Gini, data = inequality)
cor.test(GII ~ HDI, data = inequality)
cor.test(Gini ~ HDI, data = inequality)
```

#Linear model
##Research model
```{r}
GII.mod = lm(GII ~ HDI + Gini, data = inequality)
summary(GII.mod)
```
###Testing the multicollinearity
```{r}
vif(GII.mod)
```
All are below 5, so we can include all of them to the model without having concern about multicollinearity.

##Testing the model's significance
```{r}
GII.mod.reduced = lm(GII ~ 1, data = inequality)
summary(GII.mod.reduced)
```
###ANOVA
```{r}
anova(GII.mod.reduced, GII.mod)
```
Due to the above error caused by the NA values, I created another model, which dropped out the NA values, to test the significance of the model. Comparing the model summary, this dropping action's effect may be neglectable.
```{r}
ineq_nona = na.omit(inequality)
model1nona = lm(GII ~ HDI + Gini, data = ineq_nona)
summary(model1nona)
model2nona = lm(GII ~ 1, data = ineq_nona)
summary(model2nona)
anova(model2nona, model1nona)
```
##Testing the assumption
```{r}
plot(model1nona, which = 1)
plot(model1nona, which = 2)
```








